### 2022-10-22 dissertation/chapters/04-hyper-heuristics.tex
🚧 WIP - Save point (compiles)

Automation alone is not enough. It is often the case that a single configuration is not sufficient. A modern suggestion is to include a hybridisation approach where multiple configurations are considered together. This is the nature of \acp{HH}. This is different from \textit{ensemble} \cite{ref:dietterich:2002} approaches in the sense that \acp{HH} provide a single solution, where as \textit{ensembles} make use of the combination of solutions that undergo a net weighted transformation in order to produce a final solution.

@@

Meta learning is a branch of metacognition and is concerned with learning about the learning processes. Meta-learning studies how learning systems can increase in efficiency through experience and is broadly defined as the learning process concerned with the concept of \textit{learning to learn} \cite{ref:vilalta:2002}. The goal of meta-learning is to understand if learning itself can become flexible to the problem domain or task under consideration.

\citeauthor{ref:vilalta:2002} \cite{ref:vilalta:2002} mentions that meta-learning differs from base-learning in the scope of the level of adaptation. Meta-learning is concerned with learning how to choosing the right configuration and biases dynamically. On the contrary, for \textit{base}-learning, biases and configurations are predefined and fixed a priori.

@@

The term \acl{HH} was first used in 1997 \cite{ref:burke:2010} and was used to describe a protocol that combines several \ac{AI} methods in the context of automated theorem proving. The term was independently used in 2000 \cite{ref:cowling:2000} to describe ``heuristics to choose heuristics''.

Formal definition of \acp{HH} is required. \citeauthor{ref:burke:2010}\cite{ref:burke:2010} defines \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems. Burke goes on to say that \ac{HH} is a high-level approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point \cite{ref:burke2003}.

\citeauthor{ref:grobler:2015} \cite{ref:grobler:2015} states that \acp{HH} promote the design of more generally applicable search methodologies and tend to performing relatively well on a large set of different problems. This is in contrast to specialised algorithms which typically focus on outperforming the state-of-the-art for a single application.

@@

\citeauthor{ref:grobler:2015}\cite{ref:grobler:2015} mentions that two fundamental ideas behind the notion of \ac{HH} can be identified. Firstly, the recognition that the process of selecting or designing efficient hybrid and/or cooperative heuristics can be regarded as a computational search problem in itself. Secondly, there is significant potential to improve search methodologies by the incorporation of learning mechanisms that can adaptively guide the search. These two fundamental ideas have inspired different types of hyper-heuristics \cite{ref:burke:2010}.

@@

\citeauthor{ref:burke:2010}\cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the nature of the heuristic search space and the source of feedback during learning. Figure \ref{fig:heuristics:hh:classification} was taken from \cite{ref:burke:2010} and summarises the classification scheme for \ac{HH}.

@@

According to the classification proposed by \cite{ref:burke:2010}, the first dimension of classification involves the nature of the search space. Distinction is made between heuristic selection and heuristic generation. For heuristic selection \acp{HH}, methodologies are implemented for choosing or selecting existing low-level heuristics. On the contrary, for heuristic generation \acp{HH}, methodologies  are implemented for generating new heuristics from components of existing heuristics.

Selection mechanisms can include probabilistic methods as is introduced in this dissertation or even some high level \ac{MH}. For generation mechanisms, \ac{EA} are often used \cite{ref:burke:2010}.

@@

The second dimension of classification distinguishes between construction heuristics and perturbation heuristics. Consider first construction heuristics. According to \citeauthor{ref:burke:2010}\cite{ref:burke:2010}, these approaches build a solution incrementally. Starting with an empty solution, the goal is to intelligently select and use construction heuristics to gradually build a complete solution. As such, the \ac{HH} framework is provided with a set of pre-existing construction heuristics. These heuristics are generally problem specific and the challenge is to select the heuristic that is somehow the most suitable for the current problem state. This process is repeated until some final state such as a complete solution, is obtained.

@@

Heuristic selection can be done in a \textit{non-adaptive} or simple way. This is done either randomly or along a cycle and is based on a prefixed heuristic ordering \cite{ref:cowling:2000}. No learning is involved in these approaches. As an alternative, heuristic selection may incorporate an \textit{adaptive}, \textit{online learning}) mechanism based on the probabilistic weighting of the low-level heuristics \cite{ref:burke:2003} or some type of performance statistics \cite{ref:cowling:2000}. It will later be shown that the proposed \ac{BHH} implements this adaptive selection, perturbation strategy.

The acceptance strategy is an important component of any local search heuristic. \cite{ref:burke:2003} mentions that many acceptance strategies have been explored within \acp{HH}. Move acceptance strategies can be divided into two categories. These include \textit{deterministic} and \textit{non-deterministic}. In general, a move is accepted or rejected, based on the quality of the move and the current solution during a single point search. At any point in the search, deterministic move acceptance methods generate the same result for the same candidate solution(s) used for the acceptance test. However, a different outcome is possible if a non-deterministic approach is used. If the move acceptance test involves other parameters, such as the current time, then these strategies are referred to as non-deterministic strategies \cite{ref:burke:2003}.

@@

Training of \ac{FFNN} is notorious for being sensitive to configurations made beforehand. \acp{HH} provide a dynamic framework to train \ac{FFNN}. However, at the time of writing, only two examples of published work where found where \ac{HH} are used for training \ac{FFNN}. These include the works of Nel \cite{ref:nel:2021} and Ul et al. \cite{ref:ul:2018}

>>>>

Automation alone is not enough. It is often the case that a single configuration is not sufficient. A modern suggestion is to include a hybridisation approach where multiple configurations are considered together. This is the nature of \acp{HH}. This is different from \textit{ensemble}~\cite{ref:dietterich:2002} approaches in the sense that \acp{HH} provide a single solution, where as \textit{ensembles} make use of the combination of solutions that undergo a net weighted transformation in order to produce a final solution.

@@

Meta learning is a branch of metacognition and is concerned with learning about the learning processes. Meta-learning studies how learning systems can increase in efficiency through experience and is broadly defined as the learning process concerned with the concept of \textit{learning to learn}~\cite{ref:vilalta:2002}. The goal of meta-learning is to understand if learning itself can become flexible to the problem domain or task under consideration.

\citeauthor{ref:vilalta:2002}~\cite{ref:vilalta:2002} mentions that meta-learning differs from base-learning in the scope of the level of adaptation. Meta-learning is concerned with learning how to choosing the right configuration and biases dynamically. On the contrary, for \textit{base}-learning, biases and configurations are predefined and fixed a priori.

@@

The term \acl{HH} was first used in 1997~\cite{ref:burke:2010} and was used to describe a protocol that combines several \ac{AI} methods in the context of automated theorem proving. The term was independently used in 2000~\cite{ref:cowling:2000} to describe ``heuristics to choose heuristics''.

Formal definition of \acp{HH} is required.~\citeauthor{ref:burke:2010}\cite{ref:burke:2010} defines \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems. Burke goes on to say that \ac{HH} is a high-level approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point~\cite{ref:burke2003}.

\citeauthor{ref:grobler:2015}~\cite{ref:grobler:2015} states that \acp{HH} promote the design of more generally applicable search methodologies and tend to performing relatively well on a large set of different problems. This is in contrast to specialised algorithms which typically focus on outperforming the state-of-the-art for a single application.

@@

\citeauthor{ref:grobler:2015}\cite{ref:grobler:2015} mentions that two fundamental ideas behind the notion of \ac{HH} can be identified. Firstly, the recognition that the process of selecting or designing efficient hybrid and/or cooperative heuristics can be regarded as a computational search problem in itself. Secondly, there is significant potential to improve search methodologies by the incorporation of learning mechanisms that can adaptively guide the search. These two fundamental ideas have inspired different types of hyper-heuristics~\cite{ref:burke:2010}.

@@

\citeauthor{ref:burke:2010}\cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the nature of the heuristic search space and the source of feedback during learning. Figure \ref{fig:heuristics:hh:classification} was taken from~\cite{ref:burke:2010} and summarises the classification scheme for \ac{HH}.

@@

According to the classification proposed by~\cite{ref:burke:2010}, the first dimension of classification involves the nature of the search space. Distinction is made between heuristic selection and heuristic generation. For heuristic selection \acp{HH}, methodologies are implemented for choosing or selecting existing low-level heuristics. On the contrary, for heuristic generation \acp{HH}, methodologies  are implemented for generating new heuristics from components of existing heuristics.

Selection mechanisms can include probabilistic methods as is introduced in this dissertation or even some high level \ac{MH}. For generation mechanisms, \ac{EA} are often used~\cite{ref:burke:2010}.

@@

The second dimension of classification distinguishes between construction heuristics and perturbation heuristics. Consider first construction heuristics. According to~\citeauthor{ref:burke:2010}\cite{ref:burke:2010}, these approaches build a solution incrementally. Starting with an empty solution, the goal is to intelligently select and use construction heuristics to gradually build a complete solution. As such, the \ac{HH} framework is provided with a set of pre-existing construction heuristics. These heuristics are generally problem specific and the challenge is to select the heuristic that is somehow the most suitable for the current problem state. This process is repeated until some final state such as a complete solution, is obtained.

@@

Heuristic selection can be done in a \textit{non-adaptive} or simple way. This is done either randomly or along a cycle and is based on a prefixed heuristic ordering~\cite{ref:cowling:2000}. No learning is involved in these approaches. As an alternative, heuristic selection may incorporate an \textit{adaptive}, \textit{online learning}) mechanism based on the probabilistic weighting of the low-level heuristics~\cite{ref:burke:2003} or some type of performance statistics~\cite{ref:cowling:2000}. It will later be shown that the proposed \ac{BHH} implements this adaptive selection, perturbation strategy.

The acceptance strategy is an important component of any local search heuristic.~\cite{ref:burke:2003} mentions that many acceptance strategies have been explored within \acp{HH}. Move acceptance strategies can be divided into two categories. These include \textit{deterministic} and \textit{non-deterministic}. In general, a move is accepted or rejected, based on the quality of the move and the current solution during a single point search. At any point in the search, deterministic move acceptance methods generate the same result for the same candidate solution(s) used for the acceptance test. However, a different outcome is possible if a non-deterministic approach is used. If the move acceptance test involves other parameters, such as the current time, then these strategies are referred to as non-deterministic strategies~\cite{ref:burke:2003}.

@@

Training of \ac{FFNN} is notorious for being sensitive to configurations made beforehand. \acp{HH} provide a dynamic framework to train \ac{FFNN}. However, at the time of writing, only two examples of published work where found where \ac{HH} are used for training \ac{FFNN}. These include the works of Nel~\cite{ref:nel:2021} and Ul et al.~\cite{ref:ul:2018}
